\plaintitle

### 2016-04-24 paper/paper.tex

\section{Goal and Target Audience}\label{sec:goal} 
The visualization's goal can be best summarized by the following sentence: ``Why
did they win?'' The visualization focuses on exposing relations between team
performance and their player roster over several years. It allows users to find
explanations for major improvements and declines in team performance.  The
visualization allows exploration of NBA data by lay persons.  More specifically,
the visualization does not try to offer premade explanations for phenomenons
visible in the NBA data. By providing easy and intuitive access to the data, the
visualization allows users to draw their own conclusions. The visualization's
target audience are lay persons. Specifically, fans of NBA are the core of our
target audience. This means the visualization assumes most common basketball
terms are known to the audience and does not provide basic terminology
explanations.

@@

<Tell something more about the data in terms of tall/wide data
Tell something more about pre-processing steps (why?, did we change it over time?)
rfr to picture in 1st ppt sl 60>
\section{Related Work}\label{sec:literature}

@@

<Explain that this complies with the infovis seeking mantra: Overview, zoom&filter, details on demand
Bubble = overview
statistics= details on demand
>

@@

\subsection{The bubble view}
This view supposes to inform the user of the results of the NBA play-offs per 
season. A circle is used to represent a team, the size its SRS score using a '?'
distribution, a stroke to illustrate it's region or medal won during regional 
competition and curved lines to inform the user which teams played against each 
other. The user has a timeline at its disposal to scroll through time. Note that
the team can be recognized by the logo inside the circle. Our initial desing filled
the circle with the red or blue depending on the region where the teams compete. 
However, the shade of colour was darker if their final ranking was higher in their
region. We've replaced this with the logo's of the teams as we thought this is more
user friendly and the extra information of the ranking doesn't outweight.
<picture of the bubble view compared with double elimination bracket>
A more popular visualization of a play-off is the team double elimination bracket.
Although this is a more common representation, we wanted to create a more compact
version by eliminating the recurring representation of a team. We also considered 
enclosed circles and a sunburst.
<figure of 2 alternatives https://thetuftersblog.wordpress.com/2016/03/07/team-ranking-play-off-infovis/>
Although the enclosed circles is a compact representation as well, we have continued 
with our own design as we believe this view is more original and innovative. We 
didn't encouter a similar view when searching for alternatives to visualize a 
play-off or championship. 
Our initial bubble view was too compact. It connected teams that played against
each other directly. When searching for alternatives to reflect on our choice we
discovered a similar visualization in a totally different context. 
<https://source.opennews.org/en-US/articles/nyts-512-paths-white-house/>
<comparison of original bubble view with white house paths>
This visualization gives a cleaner view of the competitors for a specific team
compared to our original sketch. Hence we decided to adapt our visualization. We
added an intermediate step between two teams to clearer indicate how teams competed
to become NBA champion.
Next to the static view, the user can interact with the visualization. When he 
hoovers over a circle of a team, only the games played by that team are shown by highliting the 
competitors (their circles) and the curved lines that connect them. This view gets 
fixated when the users clicks on a team.
He/she then automatically gets forwarded to the statistics view where the bubble 
view is still represented in the right upper corner to make sure the user keeps
an overview. The statistics view will be explained in more detail in the next section.
< Hierarchy = sorting = facilitate trends of winners
position have a meaning in the bubble view
when selecting circle => zoom/more detail
Bubble principle of 
- Symmetry (whole = play-off)
- conectedness arc = games played
Pre attentive characteristics
- size of circles
- hue to highligt selections or indicate increase decline arrows next view

@@

@@

\section{Possible improvements}
As with most projects, the result is never 'finished', meaning that there are always improvements
possible but due to timing or resource constraints the haven't all been implemented. In what 
follows, we suggest a number of features that might improve the visualization:
\begin{itemize}
    \item Smooth transitions between states could improve the user experience. For example when changing
    year, a transformation of the arrow in the statistics view would help the user see a difference between years.
    At the moment the arrows change at the blink of an eye, making it difficult for us humans with a short term memory
    to make the comparison.
    \item To further support our short term memory we could also 'save' the previous state. Eg. in the field view we could
    visualize ghost shirts to indicate which players left the team or position and highlight players that joined the team. 
    This would help the user see how the team changes over time.
    \item At the moment no attention has been payed to the process & provenance of the exploration. A user thus needs to 
    memorize his action to be able to share or reproduce his actions. This could be improved by adding URL parameters. This
    way a user at least can share or save a state of an analysis. One step further is to provide a play-feature. The visitor could 
    them share his exploration and replay it. A simple implemention would ask the visitor for a start date and an end date and
    would then 'play' the changes of a particular team and characteristic over time.
\end{itemize}

@@

\subsection{Custom visualizations}
Before exploring what is out there in the d3.js world, we made our own sketches. This is how we came up with the 3 custom views that makes our visualization. When comparing our design with alternatives we could not find something that satisfied our needs. Nonetheless, they gave us inspiration to finetune our designs. Although we are satisfied with our result and we are convinced it was the right choice to reach our goal, it had some draw backs.
We noted that other teams had more freedom in exploring multiple different visualizations which enabled them to evaluate different options with trial and error. We could have prepared our data in a format that was more compatible with out of the box examples, but this would have had too much of an impact on our result.
Next to that, although we've tried to write our code as general as possible, we're not convinced this could be shared in such a way that other people building a visualization could easily re-use our work to visualize other sports.

>>>>

\section{Goal and Target Audience}\label{sec:goal} 
The visualization's goal can be best summarized by the following sentence: ``Why
did a certain NBA-team win the Championship?'' The visualization focuses on exposing relations between team
performance and their player roster over several years. It allows users to find
explanations for major improvements and declines in team performance.  The
visualization allows exploration of NBA data by lay persons.  More specifically,
the visualization does not focus on premade explanations for phenomenons
visible in the NBA data. By providing easy and intuitive access to the data, the
visualization allows users to draw their own conclusions. The visualization's
target audience are lay persons. Specifically, fans of NBA are the core of our
target audience. This means the visualization assumes most common basketball
terms are known to the audience and does not provide basic terminology
explanations. 
< We also give basic insights in why a statistic changes (arrow views)
we try to help them explore patters and trigger them for questions
Do we create awareness?
>

@@

<Tell something more about the data in terms of tall/wide data
Mention that it's a time series
Tell something more about pre-processing steps (why?, did we change it over time?)
rfr to picture in 1st ppt sl 60>
\section{Related Work}\label{sec:literature}

@@

<Explain that this complies with the infovis seeking mantra: Overview, zoom&filter, details on demand
Bubble = overview
statistics= details on demand
Talk about th organization of the 3 views + the choice for a tiled approach
- Each page is a 'tile' that we force with fixed scrolling
- Use of the keyboard is possible (space bar to scroll + arrows, arrows for timeline)
>

@@

\subsection{The bubble view}
This view supposes to inform the user of the results of the NBA play-offs per 
season. A circle is used to represent a team, the size its SRS score using a '?'
distribution, a stroke to illustrate it's region or medal won during regional 
competition and curved lines to inform the user which teams played against each 
other. The user has a timeline at its disposal to scroll through time. Note that
the team can be recognized by the logo inside the circle. Our initial desing filled
the circle with the red or blue depending on the region where the teams compete. 
However, the shade of colour was darker if their final ranking was higher in their
region. We've replaced this with the logo's of the teams as we thought this is more
user friendly and the extra information of the ranking doesn't outweight.
<picture of the bubble view compared with double elimination bracket>
A more popular visualization of a play-off is the team double elimination bracket.
Although this is a more common representation, we wanted to create a more compact
version by eliminating the recurring representation of a team. We also considered 
enclosed circles and a sunburst.
<figure of 2 alternatives https://thetuftersblog.wordpress.com/2016/03/07/team-ranking-play-off-infovis/>
Although the enclosed circles is a compact representation as well, we have continued 
with our own design as we believe this view is more original and innovative. We 
didn't encouter a similar view when searching for alternatives to visualize a 
play-off or championship. A new visualization has the advantage that it will 
trigger persons as well.
<Talk about the importance of estethics seen in the lesson>
Our initial bubble view was too compact. It connected teams that played against
each other directly. When searching for alternatives to reflect on our choice we
discovered a similar visualization in a totally different context. 
<https://source.opennews.org/en-US/articles/nyts-512-paths-white-house/>
<comparison of original bubble view with white house paths>
This visualization gives a cleaner view of the competitors for a specific team
compared to our original sketch. Hence we decided to adapt our visualization. We
added an intermediate step between two teams to clearer indicate how teams competed
to become NBA champion.
Next to the static view, the user can interact with the visualization. When he 
hoovers over a circle of a team, only the games played by that team are shown by highliting the 
competitors (their circles) and the curved lines that connect them. This view gets 
fixated when the users clicks on a team.
He/she then automatically gets forwarded to the statistics view where the bubble 
view is still represented in the right upper corner to make sure the user keeps
an overview. The statistics view will be explained in more detail in the next section.
< Hierarchy = sorting = facilitate trends of winners
position have a meaning in the bubble view
when selecting circle => zoom/more detail
Bubble principle of 
- Symmetry (whole = play-off)
- conectedness arc = games played
Pre attentive characteristics
- size of circles
- hue to highligt selections or indicate increase decline arrows next view

@@

\subsection{Improving the result}
Talk about how we improved the data-ink ration in the end..

@@

\section{Possible improvements}
As with most projects, the result is never 'finished', meaning that there are always improvements
possible but due to timing or resource constraints the haven't all been implemented. In what 
follows, we suggest a number of features that might improve the visualization:
\begin{itemize}
    \item Smooth transitions between states could improve the user experience. For example when changing
    year, a transformation of the arrow in the statistics view would help the user see a difference between years.
    At the moment the arrows change at the blink of an eye, making it difficult for us humans with a short term memory
    to make the comparison.
    \item To further support our short term memory we could also 'save' the previous state. Eg. in the field view we could
    visualize ghost shirts to indicate which players left the team or position and highlight players that joined the team. 
    This would help the user see how the team changes over time.
    \item At the moment no attention has been payed to the process & provenance of the exploration. A user thus needs to 
    memorize his action to be able to share or reproduce his actions. This could be improved by adding URL parameters. This
    way a user at least can share or save a state of an analysis. One step further is to provide a play-feature. The visitor could 
    them share his exploration and replay it. A simple implemention would ask the visitor for a start date and an end date and
    would then 'play' the changes of a particular team and characteristic over time.
    \item We could support users further in exploring for patterns by focussing more on comparisons. We could integrate a view 
    where a visitor is able to compare two teams in the same year, or in two different years.
\end{itemize}

@@

\subsection{Custom visualizations}
Before exploring what is out there in the d3.js world, we made our own sketches. This is how we came up with the 3 custom views that makes our visualization. When comparing our design with alternatives we could not find something that satisfied our needs. Nonetheless, they gave us inspiration to finetune our designs. Although we are satisfied with our result and we are convinced it was the right choice to reach our goal, it had some draw backs.
We noted that other teams had more freedom in exploring multiple different visualizations which enabled them to evaluate different options with trial and error. We could have prepared our data in a format that was more compatible with out of the box examples, but this would have had too much of an impact on our result.
<Talk about redesign>
Next to that, although we've tried to write our code as general as possible, we're not convinced this could be shared in such a way that other people building a visualization could easily re-use our work to visualize other sports.
